String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-10-20 2023-1-19/sections/introduction.tex
crv version finished

\medskip
\noindent 
\emph{Contribution.} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur (called RECL for brevity). We make the following contributions.
\begin{itemize}
  \item We propose an automata-theoretical approach for solving RECL constraints. 
  Our main idea is to represent the counting operators by cost registers in cost-enriched finite automata (CEFA, see Section~\ref{sec:automaton} for the definition), instead of unfolding them explicitly. The string-length function is modeled by cost registers as well. The satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA w.r.t. a linear integer arithmetic (LIA) formula. According to the results from~\cite{atva2020}, an LIA formula can be computed to represent the potential values of registers in CEFA.
Thus, the nonemptiness of CEFA w.r.t. LIA formulas can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers.
  \item We propose techniques to reduce the sizes (i.e. the number of states and transitions) of CEFA, in order to achieve better performance.
  \item Combined with the size-reduction techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. We implement the algorithm on top of OSTRICH, resulting in a string solver called $\ostrichrecl$. 
 %
  \item Finally, we utilize a collection of benchmark suites comprising 48,843 instances in total to evaluate the performance of $\ostrichrecl$. The experiment results show that $\ostrichrecl$ solves the RECL constraints more efficiently than the state-of-the-art string solvers, especially when the counting and length bounds are large (see Table~\ref{tab:results_regcol} and Figure~\ref{fig:table_large_count}). For instance, on 1,873 benchmark instances where the counting bounds are greater than or equal to $50$ and the string lengths are required to be beyond $200$, $\ostrichrecl$ solves at least $278$ more instances than the other solvers, while spending only half or less time per instance on average. 
\end{itemize}

>>>>

\medskip
\noindent 
\emph{Contribution.} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur (called RECL for brevity). We make the following contributions.
\begin{itemize}
  \item We propose an automata-theoretical approach for solving RECL constraints. 
  Our main idea is to represent the counting operators by cost registers in cost-enriched finite automata (CEFA, see Section~\ref{sec:automaton} for the definition), instead of unfolding them explicitly. The string-length function is modeled by cost registers as well. The satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA w.r.t. a linear integer arithmetic (LIA) formula. According to the results from~\cite{atva2020}, an LIA formula can be computed to represent the potential values of registers in CEFA.
Thus, the nonemptiness of CEFA w.r.t. LIA formulas can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers.
  \item We propose techniques to reduce the sizes (i.e. the number of states and transitions) of CEFA, in order to achieve better performance.
  \item Combined with the size-reduction techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. We implement the algorithm on top of OSTRICH, resulting in a string solver called $\ostrichrecl$. 
 %
  \item Finally, we utilize a collection of benchmark suites comprising 48,843 instances in total to evaluate the performance of $\ostrichrecl$. The experiment results show that $\ostrichrecl$ solves the RECL constraints more efficiently than the state-of-the-art string solvers, especially when the counting and length bounds are large (see Figure~\ref{fig:table_overall_eval} and Table~\ref{tab:large-bound}). For instance, on 1,969 benchmark instances where the counting bounds are greater than or equal to $50$ and the string lengths are required to be beyond $200$, $\ostrichrecl$ solves at least $278$ more instances than the other solvers, while spending only half or less time per instance on average. 
\end{itemize}
